---
title: "Task 7 - Tidyverse"
output: md_document
---
To Do:
-Tables for weather data
-Select 26-29 most common baby names

```{r}
library(tidyverse)
library(dplyr)
#comment out install.packages before knitting
#install.packages("nycflights13")
#install.packages("babynames")
```

Using the nycflights13::weather:

1) Determine whether there are any clear outliers in wind speed (wind_speed) that should be rejected. If so, filter those bad point(s) and proceed.
What direction has the highest median speed at each airport? Make a table and a plot of median wind speed by direction, for each airport. 
```{r}
nycflights13::weather %>%
  #determine outliers by graphing it
ggplot(aes(x=month, y=wind_speed)) +
  geom_point()
  #there is a clear outlier at the top, manually set filter
nycflights13::weather %>% 
  filter(wind_speed < 250) %>%
  group_by(origin, wind_dir)%>%
  summarise(med_windspeed = median(wind_speed))%>%
  #top median wind speeds per airport and direction
  top_n(1)
#table median wind speed by direction by airport
nycflights13::weather %>% 
  filter(wind_speed < 250) %>%
  group_by(origin, wind_dir)%>%
  summarise(med_windspeed = median(wind_speed))%>%
  spread(origin, med_windspeed)
#plot
nycflights13::weather %>%
  filter(wind_speed < 250) %>%
  group_by(origin, wind_dir) %>%
  summarise(med_windspeed = median(wind_speed)) %>%
  ggplot(aes(x=wind_dir, y=med_windspeed)) +
    geom_point() +
    facet_wrap(~origin)
  
```
Using nycflights13::flights and nycflights13::airlines:

2) Make a table with two columns: airline name (not carrier code) and median distance flown from JFK airport. The table should be arranged in order of decreasing mean flight distance. Hint: use a _join function to join flights and airlines.
```{r}
nycflights13::flights  %>%
  filter(origin == "JFK") %>%
  select(carrier, origin, distance) %>%
  group_by(carrier) %>%
  summarise(med_distance = median(distance)) %>%
  left_join(nycflights13::airlines, by="carrier")  %>%
  select(name, med_distance) %>%
  arrange(desc(med_distance))
```

Using nycflights13::flights and nycflights13::airlines:
3) Make a wide-format data frame that displays the number of flights that leave Newark ("EWR") airport each month, from each airline
```{r}
nycflights13::flights %>%
  filter(origin == "EWR") %>%
  select(origin, carrier, month) %>%
  group_by(carrier, month) %>%
  summarise(Flights = n()) %>%
  left_join(nycflights13::airlines, by="carrier") %>%
  select(name, month, Flights) %>%
#why does this work???????
  #spread(new columns, what to populate them with)
  spread(month, Flights)
```

Using the babynames dataset:
4) Identify the ten most common male and female names in 2014. Make a plot of their frequency (prop) since 1880. (This may require two separate piped statements).

```{r}
#identify top 10 male names in 2014
babynames::babynames %>%
  filter(sex=="M", year == "2014") %>%
  top_n(10) %>%
  select(name) %>%
#what makes left join different from right join --> not quite sure why this works but it does
  left_join(babynames::babynames, by = "name") %>%
#refilter because Noah is a M and F name
  filter(sex=="M") %>%
  ggplot(aes(x=year, y=prop)) +
    geom_point() +
    ggtitle("Frequency of 2014 Top Male Baby Names Over Time") +
    facet_wrap("name")
  
#identify top 10 female names in 2014
babynames::babynames %>%
  filter(sex == "F", year == "2014") %>%
  top_n(10) %>%
  select(name) %>%
  left_join(babynames::babynames, by = "name") %>%
  filter(sex == "F") %>%
  ggplot(aes(x=year, y=prop)) +
    geom_point() +
    ggtitle("Frequency of 2014 Top Female Baby Names Over Time") +
    facet_wrap("name")
```
Using the babynames dataset:
5) Make a single table of the 26th through 29th most common girls names in the year 1896, 1942, and 2016
```{r}
babynames::babynames %>%
  filter(sex == "F", year == "1896"| year == "1942"| year == "2016") %>%
  select(year, name, n) %>%
  group_by(year) %>%
#Figure out why this works... Select the first 29 instances based on n
  top_n(29, n) %>%
#Then from those 29, the bottom 4 rows would be 26-29
  top_n(-4, n) %>%
  print()
#no data from 2016
```